---
title: "mtcars - Transmission, Fuel efficiency and Horsepower"
author: "Max Callaghan & Gabriel Tarriba"
date: "September 30, 2015"
output: 
  ioslides_presentation:
    css: style.css
---

```{r include=FALSE}

library(ggplot2)
library(knitr)

try_user <- function(x,p) {
  a <- try(setwd(p),silent=TRUE)
}

try_user("MC","~/Documents/hertie/datascience/assignment1")
try_user("GT","/Users/Gabriel/Desktop/GabrielRepo")

source("functions.R")

```


## mtcars

We looked at the mtcars dataset (summary below). We focused on miles per gallon, automatic transmission and horsepower 

```{r echo=FALSE}
summary(mtcars)
```


## Miles per Gallon

Miles per gallon are right-skewed and the mean is `r mean(mtcars$mpg,na.rm=TRUE)` and a standard error of `r std.error(mtcars$mpg)`

```{r echo=FALSE}

hist(mtcars$mpg,breaks=10)

```


## Automatic Transmission

There are fewer automatic cars than manual cars.

```{r echo=FALSE}

ggplot(mtcars, aes(factor(am))) + geom_bar()

```

## Automatic Transmission and Miles per Gallon

A box and whisker plot seems to show that automatic cars are more efficient than manual cars

```{r echo=FALSE}

ggplot(mtcars,aes(factor(am),mpg)) + geom_boxplot()

```

## Automatic Transmission and Miles per Gallon

A Welch Two Sample t-test shows that the probability that the difference in the mean mpg between the groups manual and automatic is due to random chance is less than 1 percent. We reject the null hypothesis that the means are equal

```{r echo=FALSE}

t.test(mtcars$mpg~mtcars$am)

```

## Efficiency index

We create an index that compares the mpg in each car to the most efficient car in the sample and plot a box whisper chart by category

```{r echo=FALSE}

mtcars$eff_index <- mtcars$mpg/max(mtcars$mpg,na.rm=FALSE)*100

mtcars$eff_cat <- cut(mtcars$eff_index,c(0,25,50,75,100),labels=c("very low","low","high","very high"))

ggplot(mtcars,aes(factor(eff_cat),hp)) + geom_boxplot()

```

## Data
Our dataset now looks like this:

```{r echo=FALSE}

kable(head(mtcars))

```

