---
title: "mtcars - Transmission, Fuel Efficiency and Horsepower"
author: "Max Callaghan & Gabriel Tarriba"
date: "September 30, 2015"
output: 
  ioslides_presentation:
    css: style.css
---

```{r include=FALSE}

library(ggplot2)
library(knitr)
library(pander)

try_user <- function(x,p) {
  a <- try(setwd(p),silent=TRUE)
}

try_user("MC","~/Documents/hertie/datascience/assignment1")
try_user("GT","/Users/Gabriel/Desktop/GabrielRepo")

source("functions.R")

```


## mtcars

We looked at the mtcars dataset (summary below). We want to explore the relationship between miles per gallon, automatic transmission and horsepower. 

```{r echo=FALSE}
kable(summary(mtcars[,c("mpg","hp","am")]))
```


## Miles per Gallon

Miles per gallon are right-skewed and the mean is `r mean(mtcars$mpg,na.rm=TRUE)` and a standard error of `r std.error(mtcars$mpg)`

```{r echo=FALSE}

hist(mtcars$mpg,breaks=10,xlab="Miles per Gallon",main="Histogram of Miles per Gallon")

```


## Automatic Transmission

There are fewer automatic cars than manual cars.

```{r echo=FALSE}

ggplot(mtcars, aes(factor(am))) + 
  geom_bar() + 
  xlab("Manual (0) or Automatic (1)") + 
  theme_bw()

```

## Automatic Transmission and Miles per Gallon

A box and whisker plot seems to show that automatic cars are more fuel-efficient than manual cars

```{r echo=FALSE, fig.width=5, fig.height=4}

ggplot(mtcars,aes(factor(am),mpg)) + 
  geom_boxplot() + 
  xlab("Manual (0) or Automatic (1)") + 
  ylab("Miles per Gallon") + 
  theme_bw()

```

## Automatic Transmission and Miles per Gallon

A Welch Two Sample t-test shows that the probability that the difference in the mean mpg between the groups manual and automatic is due to random chance is less than 1 percent. We reject the null hypothesis that the means are equal. Thus we conclude that automatic cars are more fuel-efficient.

```{r echo=FALSE}

pander(t.test(mtcars$mpg~mtcars$am))

```

## Efficiency index

We create an index that compares the mpg in each car to the most efficient car in the sample. We name this variable efficiency index (in code eff_index). Then we create a categorical variable called efficiency category (in code eff_cat), which contains the categories "very low", "low", "high" and "very high" according to four quartiles the efficiency index variable

```{r echo=FALSE, fig.width=5, fig.height=4}

mtcars$eff_index <- mtcars$mpg/max(mtcars$mpg,na.rm=FALSE)*100

mtcars$eff_cat <- cut(mtcars$eff_index,c(0,25,50,75,100),labels=c("very low","low","high","very high"))

ggplot(mtcars,aes(factor(eff_cat),hp)) + geom_boxplot() +
  xlab("Efficiency Category") + 
  ylab("Horsepower") + 
  theme_bw()

```

## Data
Our dataset now looks like this:

```{r echo=FALSE}

kable(head(mtcars))

```

