---
title: "mtcars - Transmission, Fuel Efficiency and Horsepower"
author: "Max Callaghan & Gabriel Tarriba"
date: "September 30, 2015"
output: 
  ioslides_presentation:
    css: style.css
---

```{r include=FALSE}
#load some libraries
library(ggplot2)
library(knitr)
library(pander)

# define a function which quietly tries to set the working directory, without causing an error if it doesn't exist
try_user <- function(x,p) {
  a <- try(setwd(p),silent=TRUE)
}

#try to set the working directory to either Max or Gabriel's local directory
try_user("MC","~/Documents/hertie/datascience/assignment1")
try_user("GT","/Users/Gabriel/Desktop/GabrielRepo")

#run a 'dynamically linked file' which contains a function we defined
source("functions.R")

```


## mtcars

We looked at the mtcars dataset (summary below). We focused on miles per gallon, automatic transmission and horsepower 

```{r echo=FALSE}

#summarise the variables we are interested in mtcars and put them into a pretty table
kable(summary(mtcars[,c("mpg","hp","am")]))

```


## Miles per Gallon

Miles per gallon are right-skewed and the mean is `r mean(mtcars$mpg,na.rm=TRUE)` and a standard error of `r std.error(mtcars$mpg)`

```{r echo=FALSE}

#do a histogram of mpg
hist(mtcars$mpg,breaks=10,xlab="Miles per Gallon",main="Histogram of Miles per Gallon")

```


## Automatic Transmission

There are fewer automatic cars than manual cars.

```{r echo=FALSE}

#do a bar chart of transmission
ggplot(mtcars, aes(factor(am))) + 
  geom_bar() + 
  xlab("Manual (0) or Automatic (1)") + 
  theme_bw()

```

## Automatic Transmission and Miles per Gallon

A box and whisker plot seems to show that automatic cars are more efficient than manual cars

```{r echo=FALSE, fig.width=5, fig.height=4}

#do a boxplot of mpg, grouped by transmission type
ggplot(mtcars,aes(factor(am),mpg)) + 
  geom_boxplot() + 
  xlab("Manual (0) or Automatic (1)") + 
  ylab("Miles per Gallon") + 
  theme_bw()

```

## Automatic Transmission and Miles per Gallon

A Welch Two Sample t-test shows that the probability that the difference in the mean mpg between the groups manual and automatic is due to random chance is less than 1 percent. We reject the null hypothesis that the means are equal

```{r echo=FALSE}

#do a t-test of the mean of mpg grouped by transmission type and report the results in a pretty table
pander(t.test(mtcars$mpg~mtcars$am))

```

## Efficiency index

We create an index that compares the mpg in each car to the most efficient car in the sample and plot a box whisper chart by category

```{r echo=FALSE, fig.width=5, fig.height=4}

#create a new efficiency index which measures the mpg of each car as a percentage of the maximum
mtcars$eff_index <- mtcars$mpg/max(mtcars$mpg,na.rm=FALSE)*100

#create a categorical variable based on the above index
mtcars$eff_cat <- cut(mtcars$eff_index,c(0,25,50,75,100),labels=c("very low","low","high","very high"))

#do a boxplot of horsepower grouped by efficiency category
ggplot(mtcars,aes(factor(eff_cat),hp)) + geom_boxplot() +
  xlab("Efficiency Category") + 
  ylab("Horsepower") + 
  theme_bw()

```

## Data
Our dataset now looks like this:

```{r echo=FALSE}

#output the first few rows of our data in a pretty table
kable(head(mtcars))

```

