---
title: "Catholics and Fertility"
author: "Max Callaghan & Gabriel Tarriba"
date: "September 25, 2015"
output: ioslides_presentation
---

```{r include=FALSE}

# define a function which quietly tries to set the working directory, without causing an error if it doesn't exist
try_user <- function(p) {
  a <- try(setwd(p),silent=TRUE)
}

#try to set the working directory to either Max or Gabriel's local directory
try_user("~/Documents/hertie/datascience/assignment1")
try_user("/Users/Gabriel/Desktop/GabrielRepo")

#load some libraries
library(ggplot2)
library(knitr)
library(texreg)

#load the swiss dataset
data(swiss)

#run a 'dynamically linked file' which contains a function we defined
source("functions.R")

```

## Data
We are analysing the swiss datset. We want to look at the relationship between Catholicism and fertility

```{r echo=FALSE}

#summarise the swiss dataset
kable(summary(swiss))

```

## Catholicism
The distribution of Catholic density (proportion of Catholics in total population) by canton is bimodal

```{r echo=FALSE}

#do a histogram of the variable 'Catholic'
hist(swiss$Catholic,xlab="Catholic Density", main="Histogram of Catholic Density")

```

## Catholics
- We assume, given that the swiss dataset is from 1888, that the proportion of protestants in each canton is one minus the proportion of Catholics, and that swiss cantons are either mostly Catholic or mostly Protestant, with very few mixed cantons.

## Fertility
Fertility is fairly normally distributed with a mean of `r mean(swiss$Fertility,na.rm=TRUE)` and a standard error of `r std.error(swiss$Fertility)`

```{r echo=FALSE}

#do a histogram of fertility (above inline code computes mean and standard error)
hist(swiss$Fertility, xlab="Fertility Rate", main="Histogram of Fertility Rate")

```

## Relationship between Catholicism and Fertility
Plotting Fertility and Catholicism gives us strange results

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=5, fig.height=3.5}

#plot Fertility against Catholicism with points, draw a smooth line, and make the plot monochrome
ggplot(swiss, aes(Catholic,Fertility))+geom_point() + geom_smooth() + theme_bw()

```

## Relationship Religious Concentration and Fertility

We create a new variable, religious concentration, and plot it as a dependent variable, here the relationship is much more as we expected: concentration is related with fertility.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=5, fig.height=3.5}

#define religious concentration conditionally upon whether Catholic is higher or imputed Protestant (100-Catholic) is higher
swiss$religious_concentration <- ifelse(
  swiss$Catholic > 50,
  swiss$Catholic,
  100 - swiss$Catholic
)

#plot new religious concentration variable against fertility, with similar options to above
ggplot(swiss, aes(religious_concentration,Fertility)) +
  geom_point() + 
  geom_smooth() + 
  theme_bw() +
  xlab("Religious Concentration")

```


## Relationship Religious Concentration and Fertility

A simple bivariate regression suggests that there is a statistically significant positiv relationship between religious concentration and fertility

```{r echo=FALSE, results='asis', warning=FALSE}

#Do a bivariate regression where y = Fertility and x = religious_concentration
reg1 <- lm(Fertility~religious_concentration,swiss)

#report the results of the regression in a pretty table
htmlreg(reg1,single.row=TRUE)

```

